---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

**Problem1—Predicting Income using Logistic Regression and Decision Trees**

1.  Loading and Inspecting Data.

    ```{r}
    database <- read.csv("adult.data", sep = ",", na.strings = " ?", header = FALSE)
    colnames(database) <- c("age", "workclass", "fnlwgt", "education", "edunum", "marital", "occupation", "relation", "race", "sex", "capgain", "caploss", "hrsweek", "nativity", "income")
    ```

    Structure and Summary

    ```{r}
    str(database)
    summary(database)
    ```

    | Variable   | Type        | Sub-Type   |
    |------------|-------------|------------|
    | age        | Numerical   | Continuous |
    | workclass  | Categorical | Nominal    |
    | fnlwgt     | Numerical   | Continuous |
    | education  | Categorical | Ordinal    |
    | edunum     | Numerical   | Ordinal    |
    | marital    | Categorical | Nominal    |
    | occupation | Categorical | Nominal    |
    | relation   | Categorical | Nominal    |
    | race       | Categorical | Nominal    |
    | sex        | Categorical | Binary     |
    | cap-gain   | Numerical   | Continuous |
    | cap-loss   | Numerical   | Continuous |
    | hrs/week   | Numerical   | Continuous |
    | nativity   | Categorical | Nominal    |
    | income     | Categorical | Binary     |

2.  Converting ? to NA.

    while reading the data, I've used space (" ") before "?" and set those to NA.

3.  Set Seed.

    Split Data - Let's do this after plotting and doing tests - it can keep the process much clearer.

    ```{r}
    set.seed(69)
    ```

4.  Handling Missing Data

    ```{r}
    colSums(is.na(database))
    ```

    In the dataset Workclass, Occupation, Nativity has the missing values. Let's replace the missing values in those with "Unknown". The target variable is already imbalced and I believe perform mode imputation makes the machine biased, it's better to make data little diverse than making machine biased. So I did Missing Value Imputation, this enhances the statistical power of the analysis, improving the ability to detect patterns and test hypotheses effectively.

    ```{r}
    library(dplyr)

    database <- database |>
      mutate(
        workclass = ifelse(is.na(workclass), "Unknown", workclass),
        occupation = ifelse(is.na(occupation), "Unknown", occupation),
        nativity = ifelse(is.na(nativity), "Unknown", nativity)
      )
    ```

5.  Data in the variable "nativity" has various categories, let's find the frequency of the data.

    ```{r}
    library(tidyverse)
    fct_count(database$nativity, sort = TRUE, prop = TRUE)
    ```

    The most occurring country in both train_data and test_data is United States. The rest of the countries occur very least. Factoring nativity based on geography is a bit complex, so let's do the factoring based on their proportion only (top 3)

    ```{r}
    library(forcats)
    database$nativity <- fct_lump_n(database$nativity,n =3)
    table(database$nativity)
    ```

6.  Now Let's perform some plotting and tests (I factored Categorical Variables)

    ```{r echo=FALSE}
    database$workclass <- factor(database$workclass)
    database$education <- factor(database$education)
    database$marital <- factor(database$marital)
    database$occupation <- factor(database$occupation)
    database$relation <- factor(database$relation)
    database$race <- factor(database$race)
    database$sex <- factor(database$sex)
    database$nativity <- factor(database$nativity)
    database$income <- factor(database$income)
    ```

    We know that income is a categorical variable - So we will perform

    Chi Square test and Mosaic Plots on workclass, education, marital, occupation, relation, race, sex, nativity.

    ```{r}
    wo_in <- table(database$workclass,database$income)
    ed_in <- table(database$education,database$income)
    ma_in <- table(database$marital,database$income)
    oc_in <- table(database$occupation,database$income)
    re_in <- table(database$relation,database$income)
    ra_in <- table(database$race,database$income)
    se_in <- table(database$sex,database$income)
    na_in <- table(database$nativity,database$income)
    ```

    WORKCLASS

    ```{r echo=FALSE}
    mosaicplot(wo_in, main = "Mosaic Plot of income by workclass", xlab = "income", ylab = "workclass", color = TRUE)
    print(chisq.test(wo_in))
    ```

    EDUCATION

    ```{r echo=FALSE}
    mosaicplot(ed_in, main = "Mosaic Plot of income by education", xlab = "income", ylab = "education", color = TRUE)
    print(chisq.test(ed_in))
    ```

    MARITAL

    ```{r echo=FALSE}
    mosaicplot(ma_in, main = "Mosaic Plot of income by marital", xlab = "income", ylab = "marital", color = TRUE)
    print(chisq.test(ma_in))
    ```

    OCCUPATION

    ```{r echo=FALSE}
    mosaicplot(oc_in, main = "Mosaic Plot of income by occupation", xlab = "income", ylab = "occupation", color = TRUE)
    print(chisq.test(oc_in))
    ```

    RELATION

    ```{r echo=FALSE}
    mosaicplot(re_in, main = "Mosaic Plot of income by relation", xlab = "income", ylab = "relation", color = TRUE)
    print(chisq.test(re_in))
    ```

    RACE

    ```{r echo=FALSE}
    mosaicplot(ra_in, main = "Mosaic Plot of income by race", xlab = "income", ylab = "race", color = TRUE)
    print(chisq.test(ra_in))
    ```

    SEX

    ```{r echo=FALSE}
    mosaicplot(se_in, main = "Mosaic Plot of income by sex", xlab = "income", ylab = "sex", color = TRUE)
    print(chisq.test(se_in))
    ```

    NATIVITY

    ```{r echo=FALSE}
    mosaicplot(na_in, main = "Mosaic Plot of income by nativity", xlab = "income", ylab = "nativity", color = TRUE)
    print(chisq.test(na_in))
    ```

    All the above categoricals gave a low p value (i.e., p \< 2.2e-16), this implies they are significantly associated with income.

    Side-Side Box for all numeric variables, t-test for age, fnlwgt, cap-gain, cap-loss, hrs/week.

    ANOVA on edu-num.

    AGE

    ```{r echo=FALSE}
    boxplot(database$age ~ database$income, main = "Age by income", xlab = "income", ylab = "AGE")
    tt_age <- t.test(database$age ~ database$income )
    tt_age
    ```

    FNLWGT

    ```{r echo=FALSE}
    boxplot(database$fnlwgt ~ database$income, main = "fnlwgt by income", xlab = "income", ylab = "fnlwgt")
    tt_fnl <- t.test(database$fnlwgt ~ database$income )
    tt_fnl
    ```

    CAP-GAIN

    ```{r echo=FALSE}
    boxplot(database$capgain ~ database$income, main = "capital gain by income", xlab = "income", ylab = "cap-gain")
    tt_capg <- t.test(database$capgain ~ database$income )
    tt_capg
    ```

    CAP-LOSS

    ```{r echo=FALSE}
    boxplot(database$caploss ~ database$income, main = "capital loss by income", xlab = "income", ylab = "cap-loss")
    tt_capl <- t.test(database$caploss ~ database$income )
    tt_capl
    ```

    HOURS PER WEEK

    ```{r echo=FALSE}
    boxplot(database$hrsweek ~ database$income, main = "Hours per week by income", xlab = "income", ylab = "hours per week")
    tt_hw <- t.test(database$hrsweek ~ database$income )
    tt_hw
    ```

    EDUCATION NUMBER

    ```{r echo=FALSE}
    boxplot(database$edunum ~ database$income, main = "Education Number by income", xlab = "income", ylab = "edu num")
    av_edn <- aov(database$edunum ~ database$income )
    summary(av_edn)
    ```

    all the numericals are statistically sigificant to the target varible "income" except "fnlwgt".

    This implies that "fnlwgt" should be removed from the dataset.

    ```{r}
    database$fnlwgt <-NULL
    ```

    Now Splitting the data.(previously mentioned in question 3).

    ```{r}
    set.seed(532)
    train_sample <- sample((nrow(database)), floor(0.8*nrow(database)))
    train_data <- database[train_sample,]
    test_data <- database[-train_sample,]
    ```

7.  Logistic Regression

    ```{r}
    train_model <- glm(income ~ .,data = train_data, family = binomial(link = "logit"))
    summary(train_model)
    ```

    ```{r}
    predictions <- predict(train_model, test_data, type = "response")
    head(predictions)
    ```

    ```{r}
    predicted.label<-factor(ifelse(predictions>0.5, ">50k","<=50k"))
    summary(predicted.label)
    ```

8.  Cross Table - Precision and Recall

    ```{r}
    actual.label<-test_data$income
    cross_tab <- table(True = actual.label, Predicted = predicted.label)
    cross_tab
    ```

    ```{r echo=FALSE}
    tp_50 <- cross_tab[1,1]
    tn_50 <- cross_tab[2,2]    
    fp_50 <- cross_tab[2,1]   
    fn_50 <- cross_tab[1,2]
    ```

    ```{r echo=FALSE}
    tp_50k <- cross_tab[2,2]
    tn_50k <- cross_tab[1,1]
    fp_50k <- cross_tab[1,2]
    fn_50k <- cross_tab[2,1]
    ```

    We know that

    Total Error = FP + FN / (TP + TN + FP + FN)

    Precision = TP/(TP+FP)

    Recall = TP/(TP+FN)

    For income \<= 50k

    ```{r echo=FALSE}
    te_50 <- (fp_50 + fn_50)/(tp_50 + tn_50 + fp_50 + fn_50)
    pr_50 <- tp_50/(tp_50 + fp_50)
    re_50 <- tp_50/(tp_50 + fn_50)

    print(paste("Total Error =", te_50*100, "Precision =", (pr_50*100), "Recall =", re_50*100))
    ```

    For income \>50k

    ```{r echo=FALSE}
    te_50k <- (fp_50k + fn_50k)/(tp_50k + tn_50k + fp_50k + fn_50k)
    pr_50k <- tp_50k/(tp_50k + fp_50k)
    re_50k <- tp_50k/(tp_50k + fn_50k)

    print(paste("Total Error =", te_50k*100, "Precision =", (pr_50k*100), "Recall =", re_50k*100))
    ```

9.  Down-Sampling data.

    Checking Balance of the data.

    ```{r}
    table(train_data$income)
    ```

    There's an imbalance here. Let's handle it.

    ```{r}
    train_50 <- train_data[train_data$income == " <=50K",]
    train_50k <- train_data[train_data$income == " >50K",]
    ```

    ```{r}
    train_50_m <- sample_n(train_50,nrow(train_50k))
    ```

    ```{r}
    new_train_data <-rbind(train_50_m,train_50k)
    ```

    Re-Training the Data.

    ```{r}
    train_model2 <- glm(income ~ .,data = new_train_data, family = binomial(link = "logit"))
    summary(train_model2)
    ```

    ```{r}
    predictions2 <- predict(train_model2, test_data, type = "response")
    head(predictions2)
    predicted.label2<-factor(ifelse(predictions2>0.5, ">50k","<=50k"))
    summary(predicted.label2)
    cross_tab2 <- table(True = actual.label, Predicted = predicted.label2)
    cross_tab2
    ```

    ```{r echo=FALSE}
    tp_50_2 <- cross_tab2[1,1]
    tn_50_2 <- cross_tab2[2,2]    
    fp_50_2 <- cross_tab2[2,1]   
    fn_50_2 <- cross_tab2[1,2]
    tp_50k_2 <- cross_tab2[2,2]
    tn_50k_2 <- cross_tab2[1,1]
    fp_50k_2 <- cross_tab2[1,2]
    fn_50k_2 <- cross_tab2[2,1]
    ```

    For Income \<=50K

    ```{r echo=FALSE}
    te_50_2 <- (fp_50_2 + fn_50_2)/(tp_50_2 + tn_50_2 + fp_50_2 + fn_50_2)
    pr_50_2 <- tp_50_2/(tp_50_2 + fp_50_2)
    re_50_2 <- tp_50_2/(tp_50_2 + fn_50_2)

    print(paste("Total Error =", te_50_2*100, "Precision =", (pr_50_2*100), "Recall =", re_50_2*100))
    ```

    Total Error, Precision has Increased and Recall(\<=50) has Decreased when compared with full training data.

    For Income \>50K

    ```{r echo=FALSE}
    te_50k_2 <- (fp_50k_2 + fn_50k_2)/(tp_50k_2 + tn_50k_2 + fp_50k_2 + fn_50k_2)
    pr_50k_2 <- tp_50k_2/(tp_50k_2 + fp_50k_2)
    re_50k_2 <- tp_50k_2/(tp_50k_2 + fn_50k_2)

    print(paste("Total Error =", te_50k_2*100, "Precision =", (pr_50k_2*100), "Recall =", re_50k_2*100))
    ```

    Total Error, Recall has increased and Precision (\>50K)has decreased when compared with Full train data.

10. C5.0 Decision Tree Model.

    ```{r}
    library(C50)
    adult_model <- C5.0(train_data[-14], train_data$income, trials = 30)
    adult_model
    summary(adult_model)
    ```

    ```{r}
    adult_pred <- predict(adult_model, test_data)
    library(gmodels)
    CrossTable(test_data$income, adult_pred, prop.chisq = FALSE, prop.r = FALSE, prop.c = FALSE, dnn = c('actual income', 'predicted income'))
    ```

    ```{r echo=FALSE}
    adult_ctab <- table(True = actual.label, Predicted = adult_pred)
    adult_ctab

    tp_50c <- adult_ctab[1,1]
    tn_50c <- adult_ctab[2,2]    
    fp_50c <- adult_ctab[2,1]   
    fn_50c <- adult_ctab[1,2]
    tp_50kc <- adult_ctab[2,2]
    tn_50kc <- adult_ctab[1,1]
    fp_50kc <- adult_ctab[1,2]
    fn_50kc <- adult_ctab[2,1]
    ```

    For income \<=50K

    ```{r echo=FALSE}
    te_50c <- (fp_50c + fn_50c)/(tp_50c + tn_50c + fp_50c + fn_50c)
    pr_50c <- tp_50c/(tp_50c + fp_50c)
    re_50c <- tp_50c/(tp_50c + fn_50c)

    print(paste("Total Error =", te_50c*100, "Precision =", (pr_50c*100), "Recall =", re_50c*100))
    ```

    For income \>50K

    ```{r echo=FALSE}
    te_50kc <- (fp_50kc + fn_50kc)/(tp_50kc + tn_50kc + fp_50kc + fn_50kc)
    pr_50kc <- tp_50kc/(tp_50kc + fp_50kc)
    re_50kc <- tp_50kc/(tp_50kc + fn_50kc)

    print(paste("Total Error =", te_50kc*100, "Precision =", (pr_50kc*100), "Recall =", re_50kc*100))
    ```

    Comparing to Logistic Regression, C5.0 Decision tree performs better.

    Total Error is little lesser, Precision and Recall are a bit higher.

    Now C5.0 for downsized sample.

    ```{r}
    adult_model2 <- C5.0(new_train_data[-14], new_train_data$income, trials = 30)
    adult_model2
    summary(adult_model2)
    ```

    ```{r}
    adult_pred2 <- predict(adult_model2, test_data)
    library(gmodels)
    CrossTable(test_data$income, adult_pred2, prop.chisq = FALSE, prop.r = FALSE, prop.c = FALSE, dnn = c('actual income', 'predicted income'))
    ```

    ```{r echo=FALSE}
    adult_ctab2 <- table(True = actual.label, Predicted = adult_pred2)
    adult_ctab2

    tp_50c2 <- adult_ctab[1,1]
    tn_50c2 <- adult_ctab[2,2]    
    fp_50c2 <- adult_ctab[2,1]   
    fn_50c2 <- adult_ctab[1,2]
    tp_50kc2 <- adult_ctab[2,2]
    tn_50kc2 <- adult_ctab[1,1]
    fp_50kc2 <- adult_ctab[1,2]
    fn_50kc2 <- adult_ctab[2,1]
    ```

    For Income \<=50K

    ```{r echo=FALSE}
    te_50c2 <- (fp_50c2 + fn_50c2)/(tp_50c2 + tn_50c2 + fp_50c2 + fn_50c2)
    pr_50c2 <- tp_50c2/(tp_50c2 + fp_50c2)
    re_50c2 <- tp_50c2/(tp_50c2 + fn_50c2)

    print(paste("Total Error =", te_50c2*100, "Precision =", (pr_50c2*100), "Recall =", re_50c2*100))
    ```

    For Income \>50K

    ```{r echo=FALSE}
    te_50kc2 <- (fp_50kc2 + fn_50kc2)/(tp_50kc2 + tn_50kc2 + fp_50kc2 + fn_50kc2)
    pr_50kc2 <- tp_50kc2/(tp_50kc2 + fp_50kc2)
    re_50kc2 <- tp_50kc2/(tp_50kc2 + fn_50kc2)

    print(paste("Total Error =", te_50kc2*100, "Precision =", (pr_50kc2*100), "Recall =", re_50kc2*100))
    ```

    When compared the outputs of Logistic Regression and C5.0 Decision Tree Model, C5.0 performs better. i.e.., less error, higher precision and recall.

    **Problem2—Predicting Student Performance**

11. Read the dataset into a dataframe.

    ```{r}
    student <- read.csv2("student-mat.csv", sep=";", header = TRUE)
    str(student)
    summary(student)
    ```

12. Exploring the dataset.

    a\. Missing Values ?

    ```{r}
    colSums(is.na(student))
    ```

    There are no Missing Values.

    b\. Checking association.(Factoring is done in behind).

    ```{r echo = FALSE}
    student$school <- factor(student$school)
    student$sex <- factor(student$sex)
    student$address <- factor(student$address)
    student$famsize <- factor(student$famsize)
    student$Pstatus <- factor(student$Pstatus)
    student$Mjob <- factor(student$Mjob)
    student$Fjob <- factor(student$Fjob)
    student$reason <- factor(student$reason)
    student$guardian <- factor(student$guardian)
    student$schoolsup <- factor(student$schoolsup)
    student$famsup <- factor(student$famsup)
    student$paid <- factor(student$paid)
    student$activities <- factor(student$activities)
    student$nursery <- factor(student$nursery)
    student$higher <- factor(student$higher)
    student$internet <- factor(student$internet)
    student$romantic <- factor(student$romantic)
    ```

    Plotting and Testing (in loop)

    ```{r}
    factors <- colnames(student[-33])

    for (i in factors) {
      print(i)
      
      with(student, {
        if (is.factor(get(i))) {
          if (length(levels(get(i))) <= 2) {
            boxplot(G3 ~ get(i), main = paste("G3 by", i), xlab = i, ylab = "G3")
            print(t.test(G3 ~ get(i)))
          } else {
            boxplot(G3 ~ get(i), main = paste("G3 by", i), xlab = i, ylab = "G3")
            aov_result <- aov(G3 ~ get(i))
            print(summary(aov_result))
          }
        } else if (i %in% c("Fedu", "Medu")) {
          plot(G3 ~ get(i),main = paste("G3 by", i), xlab = i, ylab = "G3")
          sp_cor <- cor(G3, get(i), method = "spearman")
          print(paste("Spearman-Cor-Coefficient:", sp_cor))
        } else {
          plot(G3 ~ get(i),main = paste("G3 by", i), xlab = i, ylab = "G3")
          pr_cor <- cor(G3, get(i), method = "pearson")
          print(paste("Pearson-Cor-Coefficient:", pr_cor))
        }
      })
    }
    ```

    Keeping the Signifance level(0.05), from the above results we can say that - sex, address, Mjob, failures, schoolsup, paid, higher, internet, romantic, G1, G2 are associated with target variable G3.

    c\. Histogram of G3

    ```{r}
    hist(student$G3, main = "Histogram of G3", xlab = "G3", ylab = "Frequency")
    ```

    We can see that most students have score in the range [8,12] and I believe average(mean) is in this range, the count of people who have low score is greater than the count of people who have high score.

13. Splitting the data.

    ```{r}
    student_sample <- sample((nrow(student)), floor(0.8*nrow(student)))
    student_train <- student[student_sample,]
    student_test <- student[-student_sample,]
    ```

14. Random Seed

    ```{r}
    set.seed(123)
    ```

15. 10 fold cross validation using linear regression

    ```{r}
    library(caret)
    train.control = trainControl(method = "cv", number = 10)
    stumodel<- train(G3 ~ ., data = student_train, method = "lm", trControl = train.control)
    print(stumodel)
    summary(stumodel)
    ```

    The coefficients that are statistically different from zero are famrel, Walc, absences, G1 and G2. This means that Score in G3 somehow depend on them.

    10 fold cross validation using step wise linear regression method with backward selection.

    ```{r}
    library(leaps)
    final.model <- train(G3 ~., data = student_train, method = "leapBackward", trControl = train.control, tuneGrid = data.frame(nvmax=1:32))
    print(final.model)
    ```

    I took nvmax = 32 as there are 32 variables.

    ```{r}
    summary(final.model)
    ```

16. 10 fold cross validation using step wise linear regression method with backward selection gives lowest RSME.

    Predictions

    ```{r}
    pred <- predict(final.model, student_test)
    summary(student_test$G3)
    summary(pred)
    sqrt(mean((student_test$G3 - pred)^2))
    ```

    The first summary is of Actual test data and latter is of Predicted data. We can see that Model perform well from Minimum to Maximum.

Thank you.
